### 2020-12-19 Thesis-Chapter6-ThreatModel.tex
Editorial tweaks

\modelitem{Insider threat}{Insider threats involve attackers with roles inside the \ac{EA} recovery process who seek to
abuse the system for their own purposes. They could work in law enforcement, a device vendor, or a digital service
provider. They may have the same motivations and skills of any other criminal category, but they have the advantage of
privileged access at some stage in the process.}

@@

\modelitem{The platform abuser}{The device user him or herself may actually be a threat in the larger safety context.
Therefore, the end user's will is not the platform's top priority.\\

All technologies have embedded values \cite{rogaway_moral_2015}, and at the top of each value system is one top
priority. Traditional device security threat models prioritize the will of end user. This leads to robust design
decisions and honors the user's rights and autonomy.\\

It is dangerous to stray from this rule, but many systems already do. Auto-updating software values the vendor's update
policy over the user's preferences. Digital Rights Management software values corporations' data rights over the user's
(sometimes to extremes \cite{eff_2020}). Unencrypted e-mail software values the e-mail provider's ability to collect
personal information over the user's privacy. \Acl{EA} software values lawful investigation over the user's absolute
privacy. Intentional subversion of the user's will violates a security practitioner's instincts, but it is a frequent
reality.\\

In the broader threat model, the platform abuser uses encryption to hide illicit activities. He or she is a threat to
public safety.}

\modelitem{Hawkish lawmakers}{In the pursuit of security and privacy for society as a whole, aggressive lawmakers that
have the power to \ii{mandate} arbitrary data access present perhaps the largest threat of all. Crypto-anarchists will
encrypt their data above the compromised layer regardless, but above-the-table organizations and most individuals who
use default products will abide by U.S. law.\\

Hawkish lawmakers threaten the security of a cryptosystem (whether it uses \ac{EA} or not) because they may judge the
system as too inhibitive to law enforcement. Therefore, they may outlaw its use and mandate something weaker in its
place.}

@@

\modelitem{\ac{EA} workarounds}{The user can evade \ac{EA} mechanisms by encrypting data at a higher level in the tech
stack, for example by encrypting files before saving them to disk. Additionally, investigators or attackers can often
bypass current lockout mechanisms. This has already been discussed, and is not relevant when evaluating the \ac{EA}
system itself.}

\modelitem{Supply chain attacks}{It is necessary to protecting device hardware and software supply chains in order to
prevent surreptitious, trivially exploitable backdoors. For the purposes of this threat model, device hardware and
software are assumed to be uncompromised, aside from the standard amount of bugs and side-channels.}

@@

The \ac{UID} is not depicted in key storage because it is burned into the secure processor silicon and cannot be read by
software. The Volume key is encrypted by the Class key before it is stored. When a user updates his \ac{PIN}, the
encrypted Volume key in storage is replaced with the copy encrypted with the new Class key. Application processors do
not handle any keys or secret information besides the \ac{PIN}. Due to inline encryption in the Encryption Module,
persistent storage does not handle \ac{plaintext} data.

@@

The next step in threat modeling is to elicit threats based on the system's goals, threat actors, scope, and design.
Recall from \mysec{sec-crypto-basics} that security can be expressed in terms of authentication, integrity,
non-repudiation, confidentiality, availability, and authorization. Threats are often expressed as the opposite of these
properties: spoofing, tampering, repudiation, information disclosure, \acl{DOS}, and elevation of privileged
\cite{shostack_threat_2014}. This section analyzes \ldawmsr according to these categories. Threats that could fall into
multiple categories, such as tampering for the effect of repudiation, are listed only once.

@@

\ii{S1}, \ii{S2}, \ii{S3}, \ii{S4}, \ii{S5}, \ii{T2}, \ii{T4}, \ii{R1}, and \ii{I2} are all higher-risk threats to the
security, trustworthiness, and effectiveness of the proposal. However, they can be mitigated through a set of related
requirements: require destructive disassembly of the device to access the \ac{EA} hardware interface, require the device
to be physically present on LEA premises the entire duration it is unlocked, and grant only a handful of LE facilities
the lawful right to use the \ac{EA} mechanism. Requiring destructive disassembly entirely precludes certain attacks.
Requiring the device to be on LEA premises prevents attackers from manipulating LE into achieving the attackers' ends.
Limiting the number of LE facilities that can participate in the \ac{EA} process prevents LE impersonation attacks.
Together, these mitigations make the proposal more resistant to attacks.

@@

Finally, there are the threats of \ii{I4}, \ii{I5}, and \ii{E3}, which each represent the disclosure of unlock artifact
data (the DID-DSK database). These threats create a large amount of risk that cannot be safely mitigated. Storing so
much sensitive data in a central database creates a high-value target for foreign intelligence and cyber criminals. The
vendor needs to protect this database from the most advanced attackers in the world while also regularly updating and
accessing it. \Acp{HSM} and robust security policies would make that a difficult task. The data will almost certainly
leak eventually. The effects of such an event would depend on who stole the unlock artifacts. Criminals would use the
data for personal and financial gain; foreign intelligence and authoritarian regimes would use it to spy on their
enemies. In the worst-case scenario, anyone with specialized tools and disregard for the law could get into any
compliant device after disassembling it and waiting through the time vault.

@@

\paragraph*{Law enforcement utility} \ldawmsr includes a few tradeoffs between security and law enforcement utility. The
requirements to have physical possession of the device, to wait through the time vault, and to limit the number of
authorized law enforcement laboratories all create barriers to usability. The physical possession requirement is too
important for security to be compromised for utility. Additionally, search and seizure law is already designed for
physical evidence; maintaining the possession requirement makes the system operate within well-understood norms. The
time vault prevents against time-sensitive attacks, though destructive disassembly (not part of the original proposal)
may be enough to cover these cases. The time vault period could be decreased without significant impact. Limiting the
number of laboratories that can use the system reduces the risk of spoofing attacks; local law enforcement would have to
cooperate with one of these special laboratories, such as a regional \ac{FBI} office, to unlock a device. In the end,
law enforcement still has a reliable way to get into lawfully seized devices. \ldawmsr upholds its end, but the
government must fund the \ac{FBI} operation and cultivate cooperation between local and federal agencies.

\paragraph*{Economic impact} Expenses to the vendor fall into three categories---the costs of implementing the design,
running the system, and absorbing lost business. Even if a proposal like \ldawmsr was refined enough to be implemented,
the vendor would need to update hardware designs and create the software systems to get it started. Once in use, the
vendor would face considerable costs in running it securely due to the difficulty of safely accessing and updating the
unlock artifacts. Finally, the vendor would certainly lose privacy-conscious customers to small businesses, foreign
competitors, or resellers offering pre-\ac{EA} devices. The scale of these costs relative to device manufacturer's large
size is unclear. Costs to the government include the \ac{FBI} laboratory cost and the cost of law enforcement resources
and training. \acs{CALEA} includes provisions for government reimbursement of vendor compliance costs, and a similar
provision could be applied to a \ac{DAR} \ac{EA} regulation \cite{edwards_hr4922_1994}.

>>>>

\modelitem{Insider threat}{Insider threats are attackers with roles inside the \ac{EA} recovery process who seek to
abuse the system for their own purposes. They could work in law enforcement, a device vendor, or a digital service
provider. They may have the same motivations and skills of any other criminal category, but they have the advantage of
privileged access at some stage in the process.}

@@

\modelitem{The platform abuser}{The device user him or herself may actually be a threat in the larger safety context.
Therefore, the end user's will is not the platform's top priority.

\parboxindent All technologies have embedded values \cite{rogaway_moral_2015}, and at the top of each value system is
one top priority. Traditional device security threat models prioritize the will of end user. This leads to robust design
decisions and honors the user's rights and autonomy.

\parboxindent It is dangerous to stray from this rule, but many systems already do. Auto-updating software values the
vendor's update policy over the user's preferences. Digital Rights Management software values corporations' data rights
over the user's (sometimes to extremes \cite{eff_2020}). Unencrypted e-mail software values the e-mail provider's
ability to collect personal information over the user's privacy. \Acl{EA} systems value lawful investigation over the
user's absolute privacy. Intentional subversion of the user's will violates a security practitioner's instincts, but it
is a frequent reality.

\parboxindent In the broader threat model, the platform abuser uses encryption to hide illicit activities. He or she is
a threat to public safety.}

\modelitem{Hawkish lawmakers}{In the pursuit of security and privacy for society as a whole, aggressive lawmakers that
have the power to \ii{mandate} arbitrary data access present perhaps the largest threat of all. Crypto-anarchists will
encrypt their data above the compromised layer regardless, but above-the-table organizations and most individuals who
use default products will abide by U.S. law.

\parboxindent Hawkish lawmakers threaten the security of a cryptosystem (whether it uses \ac{EA} or not) because they
may judge the system as too inhibitive to law enforcement. Therefore, they may outlaw its use and mandate something
weaker in its place.}

@@

\modelitem{\ac{EA} workarounds}{The user can evade \ac{EA} mechanisms by encrypting data at a higher level in the tech
stack, for example by encrypting files in software before saving them to disk. Additionally, investigators or attackers
can often bypass current lockout mechanisms. This has already been discussed, and is not relevant when evaluating the
\ac{EA} system itself.}

\modelitem{Supply chain attacks}{It is necessary to protect device hardware and software supply chains in order to
prevent surreptitious, trivially exploitable backdoors. For the purposes of this threat model, device hardware and
software are assumed to be uncompromised, aside from the standard amount of bugs and side-channels.}

@@

The \ac{UID} is not depicted in key storage because it is burned into the secure processor silicon and cannot be read by
software. The Volume key is encrypted by the Class key before it is stored. When a user updates his or her \ac{PIN}, the
encrypted Volume key in storage is replaced with the copy encrypted with the new Class key. Application processors do
not handle any keys or secret information besides the \ac{PIN}. Due to inline encryption in the Encryption Module,
persistent storage does not handle \ac{plaintext} data.

@@

The next step in threat modeling is to elicit threats based on the system's goals, threat actors, scope, and design.
Recall from \mysec{sec-crypto-basics} that security can be expressed in terms of authentication, integrity,
non-repudiation, confidentiality, availability, and authorization. Threats are often expressed as the opposite of these
properties: spoofing, tampering, repudiation, information disclosure, \acl{DOS}, and elevation of privilege
\cite{shostack_threat_2014}. This section analyzes \ldawmsr according to these categories. Threats that could fall into
multiple categories, such as tampering for the effect of repudiation, are listed only once.

@@

\ii{S1}, \ii{S2}, \ii{S3}, \ii{S4}, \ii{S5}, \ii{T2}, \ii{T4}, \ii{R1}, and \ii{I2} are all higher-risk threats to the
security, trustworthiness, and effectiveness of the proposal. However, they can be mitigated through a set of related
requirements: require destructive disassembly of the device to access the \ac{EA} hardware interface, require the device
to be physically present on LEA premises the entire duration it is unlocked, and grant only a handful of LE facilities
the lawful right to use the \ac{EA} mechanism. Requiring destructive disassembly entirely precludes certain attacks.
Requiring the device to be on LEA premises prevents attackers from manipulating LE into achieving the attackers' ends.
Limiting the number of LE facilities that can participate in the \ac{EA} process prevents LE impersonation attacks.
Together, these mitigations make the proposal more resistant to attack.

@@

Finally, there are the threats of \ii{I4}, \ii{I5}, and \ii{E3}, which each represent the disclosure of unlock artifact
data (the DID-DSK database). These threats create a large amount of risk that cannot be safely mitigated. Storing so
much sensitive data in a central database creates a high-value target for foreign intelligence and cyber criminals. The
vendor needs to protect this database from the most advanced attackers in the world while also regularly updating and
accessing it. \Acp{HSM} and robust security policies would make that a difficult task. However, the data will almost
certainly leak eventually. The effects of such an event would depend on who stole the unlock artifacts. Criminals would
use the data for personal and financial gain; foreign intelligence and authoritarian regimes would use it to spy on
their enemies. In the worst-case scenario, anyone with specialized tools and disregard for the law could get into any
compliant device after disassembling it and waiting through the time vault.

@@

\paragraph*{Law enforcement utility} \ldawmsr includes a few tradeoffs between security and law enforcement utility. The
requirements to have physical possession of the device, to wait through the time vault, and to limit the number of
authorized law enforcement laboratories all create barriers to usability. The physical possession requirement is too
important for security to be compromised for utility. Additionally, search and seizure law is already designed for
physical evidence; maintaining the possession requirement makes the system operate within well-understood norms. The
time vault prevents against time-sensitive attacks, though destructive disassembly (not part of the original proposal)
may be enough to cover these cases. The time vault period could be decreased without significant impact. Limiting the
number of laboratories that can use the system reduces the risk of spoofing attacks; local law enforcement would have to
cooperate with one of these special laboratories, such as a regional \ac{FBI} office, to unlock a device. In the end,
law enforcement still has a reliable way to get into lawfully seized devices. \ldawmsr does its part, but the government
must fund the \ac{FBI} operation and cultivate cooperation between local and federal agencies.

\paragraph*{Economic impact} Expenses to the vendor fall into three categories---the costs of implementing the design,
running the system, and absorbing lost business. Even if a proposal like \ldawmsr was refined enough to be implemented,
the vendor would need to update hardware designs and create the software systems to get it started. Once in use, the
vendor would face considerable costs in running it securely due to the difficulty of safely accessing and updating the
unlock artifacts. Finally, the vendor would certainly lose privacy-conscious customers to small businesses, foreign
competitors, or resellers offering pre-\ac{EA} devices. The scale of these costs relative to device manufacturer's large
size is unclear. Costs to the government include the \ac{FBI} laboratory cost and the cost of law enforcement resources
and training. \acs{CALEA} includes provisions for government reimbursement of vendor compliance costs, and a similar
provision could be applied to \ac{DAR} \ac{EA} regulation \cite{edwards_hr4922_1994}.

### 2020-12-19 Thesis-Glossary.tex
Fix LEAD->LAED act term
Address more todos
Editorial tweaks

@@

% Start acronyms
\newabbreviation{AECA}{AECA}{Arms Export Control Act}
\newabbreviation{APT}{APT}{advanced persistent threat}
\newabbreviation{AV}{AV}{anti-virus}
\newabbreviation{AWA}{AWA}{All Writs Act}
\newabbreviation{CALEA}{CALEA}{Communications Assistance for Law Enforcement Act}
\newabbreviation{CESA}{CESA}{Cyberspace Electronic Security Act}
\newabbreviation{CFAA}{CFAA}{Computer Fraud and Abuse Act}
\newabbreviation{CIA}{CIA}{Central Intelligence Agency}
\newabbreviation{CP}{CP}{child pornography}
\newabbreviation{CSAM}{CSAM}{child sexual abuse material}
\newabbreviation{CSAV}{CSAV}{computer supported argument visualization}
\newabbreviation{DAR}{DAR}{data at rest}
\newabbreviation{DE}{DE}{device encryption}
\newabbreviation{DFD}{DFD}{data flow diagram}
\newabbreviation{DID}{DID}{Device ID}
\newabbreviation{DIM}{DIM}{data in motion}
\newabbreviation{DOS}{DOS}{denial of service}
\newabbreviation{DSK}{DSK}{Device Seal Key}
\newabbreviation{E2E}{E2E}{end to end}
\newabbreviation{E2EE}{E2EE}{end to end encryption}
\newabbreviation{EA}{EA}{exceptional access}
\newabbreviation{ECPA}{ECPA}{Electronic Communications Privacy Act}
\newabbreviation{ED}{ED}{ephemeral data}
\newabbreviation{FBI}{FBI}{Federal Bureau of Investigation}
\newabbreviation{FISA}{FISA}{Foreign Intelligence Surveillance Act}
\newabbreviation{FISAAA}{FISAAA}{FISA Amendments Act}
\newabbreviation{FISC}{FISC}{Foreign Intelligence Surveillance Court}
\newabbreviation{GCHQ}{GCHQ}{Government Communications Headquarters (UK)}
\newabbreviation{HSM}{HSM}{Hardware Security Module}
\newabbreviation{IBIS}{IBIS}{issue-based information system}
\newabbreviation{ITAR}{ITAR}{International Traffic in Arms Regulations}
\newabbreviation{KEK}{KEK}{key encrypting key}
\newabbreviation{LE}{LE}{law enforcement}
\newabbreviation{LEA}{LEA}{law enforcement agency}
\newabbreviation{LEO}{LEO}{law enforcement officer}
\newabbreviation{NED}{NED}{non-ephemeral data}
\newabbreviation{NSA}{NSA}{National Security Agency}
\newabbreviation{NSL}{NSL}{National Security Letter}
\newabbreviation{PAA}{PAA}{Protect America Act}
\newabbreviation{PGP}{PGP}{Pretty Good Privacy}
\newabbreviation{PIN}{PIN}{personal identification number}
\newabbreviation{PKI}{PKI}{public key infrastructure}
\newabbreviation{PW}{PW}{password}
\newabbreviation{SE}{SE}{strong encryption}
\newabbreviation{SSO}{SSO}{single sign-on}
\newabbreviation{TLS}{TLS}{Transport Layer Security}
\newabbreviation{UID}{UID}{Unique ID}
\newabbreviation{VPN}{VPN}{virtual private network}
% End acronyms

@@

\newabbreviation{USAPATRIOTAct}{USA PATRIOT Act}{Uniting and Strengthening America by Providing Appropriate Tools Required to Intercept and Obstruct Terrorism Act}
\newabbreviation{USAFREEDOMAct}{USA FREEDOM Act}{Uniting and Strengthening America by Fulfilling Rights and Ensuring Effective Discipline Over Monitoring Act}
\newabbreviation{SAFEAct}{SAFE Act}{Security and Freedom Through Encryption Act}
\newabbreviation{PROCODEAct}{Pro-C0DE Act}{Promotion of Commerce On-Line in the Digital Era Act}
\newabbreviation{ENCRYPTAct}{ENCRYPT Act}{Ensuring National Constitutional Rights for Your Private Telecommunications Act}
\newabbreviation{EARNITAct}{EARN IT Act}{Eliminating Abusive and Rampant Neglect of Interactive Technologies Act}
\newabbreviation{LEADAct}{LEAD Act}{Lawful Access to Encrypted Data Act}
\newabbreviation{OODAloop}{OODA Loop}{\ii{Observe-Orient-Decide-Act} Loop}

>>>>

\newterm{masssurv}{mass surveillance}{
    Ubiquitous data collection and analysis by remote, centralized bodies
}

@@

% Start acronyms
\newabbreviation{AECA}{AECA}{Arms Export Control Act}
\newabbreviation{APT}{APT}{advanced persistent threat}
\newabbreviation{AV}{AV}{anti-virus}
\newabbreviation{AWA}{AWA}{All Writs Act}
\newabbreviation{CALEA}{CALEA}{Communications Assistance for Law Enforcement Act}
\newabbreviation{CESA}{CESA}{Cyberspace Electronic Security Act}
\newabbreviation{CFAA}{CFAA}{Computer Fraud and Abuse Act}
\newabbreviation{CIA}{CIA}{Central Intelligence Agency}
\newabbreviation{CP}{CP}{child pornography}
\newabbreviation{CSAM}{CSAM}{child sexual abuse material}
\newabbreviation{CSAV}{CSAV}{computer supported argument visualization}
\newabbreviation{DAR}{DAR}{data at rest}
\newabbreviation{DE}{DE}{device encryption}
\newabbreviation{DFD}{DFD}{data flow diagram}
\newabbreviation{DID}{DID}{Device ID}
\newabbreviation{DIM}{DIM}{data in motion}
\newabbreviation{DOS}{DOS}{denial of service}
\newabbreviation{DSK}{DSK}{Device Seal Key}
\newabbreviation{E2E}{E2E}{end to end}
\newabbreviation{E2EE}{E2EE}{end to end encryption}
\newabbreviation{EA}{EA}{exceptional access}
\newabbreviation{ECPA}{ECPA}{Electronic Communications Privacy Act}
\newabbreviation{ED}{ED}{ephemeral data}
\newabbreviation{FBI}{FBI}{Federal Bureau of Investigation}
\newabbreviation{FISA}{FISA}{Foreign Intelligence Surveillance Act}
\newabbreviation{FISAAA}{FISAAA}{FISA Amendments Act}
\newabbreviation{FISC}{FISC}{Foreign Intelligence Surveillance Court}
\newabbreviation{GCHQ}{GCHQ}{Government Communications Headquarters}
\newabbreviation{HSM}{HSM}{Hardware Security Module}
\newabbreviation{IBIS}{IBIS}{issue-based information system}
\newabbreviation{ITAR}{ITAR}{International Traffic in Arms Regulations}
\newabbreviation{KEK}{KEK}{key encrypting key}
\newabbreviation{LE}{LE}{law enforcement}
\newabbreviation{LEA}{LEA}{law enforcement agency}
\newabbreviation{LEO}{LEO}{law enforcement officer}
\newabbreviation{NED}{NED}{non-ephemeral data}
\newabbreviation{NSA}{NSA}{National Security Agency}
\newabbreviation{NSL}{NSL}{National Security Letter}
\newabbreviation{PAA}{PAA}{Protect America Act}
\newabbreviation{PGP}{PGP}{Pretty Good Privacy}
\newabbreviation{PIN}{PIN}{personal identification number}
\newabbreviation{PKI}{PKI}{public key infrastructure}
\newabbreviation{PW}{PW}{password}
\newabbreviation{SE}{SE}{strong encryption}
\newabbreviation{SSO}{SSO}{single sign-on}
\newabbreviation{TLS}{TLS}{Transport Layer Security}
\newabbreviation{UID}{UID}{Unique ID}
\newabbreviation{VPN}{VPN}{virtual private network}
% End acronyms

@@

\newabbreviation{USAPATRIOTAct}{USA PATRIOT Act}{Uniting and Strengthening America by Providing Appropriate Tools Required to Intercept and Obstruct Terrorism Act}
\newabbreviation{USAFREEDOMAct}{USA FREEDOM Act}{Uniting and Strengthening America by Fulfilling Rights and Ensuring Effective Discipline Over Monitoring Act}
\newabbreviation{SAFEAct}{SAFE Act}{Security and Freedom Through Encryption Act}
\newabbreviation{PROCODEAct}{Pro-C0DE Act}{Promotion of Commerce On-Line in the Digital Era Act}
\newabbreviation{ENCRYPTAct}{ENCRYPT Act}{Ensuring National Constitutional Rights for Your Private Telecommunications Act}
\newabbreviation{EARNITAct}{EARN IT Act}{Eliminating Abusive and Rampant Neglect of Interactive Technologies Act}
\newabbreviation{LAEDAct}{LAED Act}{Lawful Access to Encrypted Data Act}
\newabbreviation{OODAloop}{OODA Loop}{\ii{Observe-Orient-Decide-Act} Loop}
\newabbreviation{SN}{S/N}{serial number}
